### 2024-11-19 README.Rmd
update READEME

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

<!-- badges: start -->
<!-- badges: end -->

# hatchR: Predicting Fish Developmental Phenology

**IN DEVELOPMENT**

`hatchR` is an R package for predicting fish developmental phenology using statistical models. It is intended to bridge the analytic gap of taking statistical models developed in hatchery settings (e.g., Beacham and Murray 1990) and applying them to real world temperature data sets using the effective value framework developed by Sparks et al. (2019).

## Citation

If you use `hatchR` in a formal publication or report, please cite it. Citing `hatchR` lets us devote more resources to it in the future. View the `hatchR` citation by running
```r
citation(package = "hatchR")
```

## Installation Instructions

You can install the development version of `hatchR` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("bmait101/hatchR")
```

## Contributing to `hatchR`

We encourage users to report bugs and/or contribute to `hatchR`. For more detail on how to do this, please see our contributing guide (`CONTRIBUTING.md`).

## Getting Help

There are several ways to get help with `hatchR`:

1. Open a GitHub issue [link here](https://github.com/bmait101/hatchR/issues).
2. Email the hatchR support team ([Morgan Sparks](morgan.sparks@usda.gov), [Bryan Maitland](bryan.maitland@usda.gov))
3. Post on a support website like Stack Overflow or Cross Validated.

## Example Usage

To come...

## Imported Packages

`hacthR` imports the following **R** packages:

* dplyr: For data manipulation.
* ggplot2: For vignette visualizations.
* lubridate: For working with dates.
* tibble: For creating tibbles as output for various functions.
* modelr: For modeling functions.
* rlang: For `.data` usage.
* stats: For various modeling functions (e.g., `confint()`).
* utils: For various utility functions.

## Suggested Packages

`hatchR` suggests the following **R** packages:

* knitr: For vignette building.
* readr: For reading in data.
* rmarkdown: For vignette building.
* testthat: For unit testing.

## License

This project is licensed under the GNU General Public License, [GPL-3](https://cran.r-project.org/web/licenses/GPL-3).

## USDA Disclaimer

`hatchR` project code is provided on an "as is" basis and the user assumes responsibility for its use. USDA Forest Service has relinquished control of the information and no longer has responsibility to protect the integrity , confidentiality, or availability of the information. Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by USDA Forest Service.