### 2024-11-22 README.Rmd
add useage to READEME

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# hatchR <a href="https://bmait101.github.io/hatchR/"><img src="man/figures/logo.png" align="right" height="139" alt="hatchR website" /></a>

<!-- badges: start -->
[![R-CMD-check](https://github.com/bmait101/hatchR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bmait101/hatchR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**IN DEVELOPMENT**

`hatchR` is an R package for predicting fish developmental phenology using statistical models. It is intended to bridge the analytic gap of taking statistical models developed in hatchery settings (e.g., Beacham and Murray 1990) and applying them to real world temperature data sets using the effective value framework developed by Sparks et al. (2019).

## Citation

If you use `hatchR` in a formal publication or report, please cite it. Citing `hatchR` lets us devote more resources to it in the future. View the `hatchR` citation by running
```r
citation(package = "hatchR")
```

## Installation Instructions

You can install the development version of `hatchR` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("bmait101/hatchR")
```

## Contributing to `hatchR`

We encourage users to report bugs and/or contribute to `hatchR`. For more detail on how to do this, please see our contributing guide (`CONTRIBUTING.md`).

## Getting Help

There are several ways to get help with `hatchR`:

1. Open a GitHub issue [link here](https://github.com/bmait101/hatchR/issues).
2. Email the hatchR support team ([Morgan Sparks](morgan.sparks@usda.gov), [Bryan Maitland](bryan.maitland@usda.gov))
3. Post on a support website like Stack Overflow or Cross Validated.

## Example Usage

Below we provide a brief example showing how to use `hatchR`. For a thorough introduction to the software, see our introductory vignette [linked here](https://bmait101.github.io/hatchR/articles/Introduction.html). For a list of all functions available in `hatchR`, see our function reference [linked here](https://bmait101.github.io/hatchR/reference/index.html).

First we load `hatchR` and helper packages by running

```{r}
library(hatchR)
library(ggplot2)  # for additional plotting options
library(lubridate)  # for working with dates
```

The example `woody_island` dataset is included in `hatchR` and contains temperature data from Woody Island, Lake Iliamna, Alaska. It includes daily water temperature data from 1990-1992. We can use `plot_check_temp()` to visually check the data:

```{r}
p <- plot_check_temp(
  data = woody_island,
  dates = date, 
  temperature = temp_c
  )
p
```

Spawning in this system typically peaks around August 18 and hatching and emergence are done before the start of the following spawning season, so we can predict phenology within a subset of a year.

```{r}
p + 
  geom_rect(
    aes(
      xmin = ymd("1990-08-18"),  # spawn date
      xmax = ymd("1991-04-01"),  # approx phenology window end
      ymin = -10,  # lower bound
      ymax = 25),  # upper bound
    fill = "darkgreen",  # color rectangle dark green
    alpha = 0.01  # make rectangle semi-transparent
    ) 
```

Next, select hatch and emerge models we want to use to predict phenology:

```{r}
sockeye_hatch_mod <- model_select(
  author = "Beacham and Murray 1990", 
  species = "sockeye", 
  model = 2, 
  dev.type = "hatch"
  )
```

Now we can predict phenology for the sockeye salmon using the `predict_phenology()` function:

```{r}
WI_hatch <- predict_phenology(
  data = woody_island,
  dates = date,
  temperature = temp_c,
  spawn.date = "1990-08-18",
  model = sockeye_hatch_mod
  )
```

We can check the predicted phenology by running:

```{r}
# see days to hatch
WI_hatch$days2done
# and development period
WI_hatch$dev.period
```

Finally, we can visualize the predicted phenology using `plot_phenology()`:

```{r}
plot_phenology(WI_hatch)
```

## Imported Packages

`hacthR` imports the following **R** packages:

* dplyr: For data manipulation.
* ggplot2: For vignette visualizations.
* lubridate: For working with dates.
* tibble: For creating tibbles as output for various functions.
* rlang: For `.data` usage.
* stats: For various modeling functions (e.g., `confint()`).
* utils: For various utility functions.

## Suggested Packages

`hatchR` suggests the following **R** packages:

* knitr: For vignette building.
* readr: For reading in data.
* rmarkdown: For vignette building.
* testthat: For unit testing.

## License

This project is licensed under the GNU General Public License, [GPL-3](https://cran.r-project.org/web/licenses/GPL-3).

## USDA Disclaimer

`hatchR` project code is provided on an "as is" basis and the user assumes responsibility for its use. USDA Forest Service has relinquished control of the information and no longer has responsibility to protect the integrity , confidentiality, or availability of the information. Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by USDA Forest Service.