### 2025-02-14 inst/manuscript/oxford/oxford.Rmd

---
title: Template for Oxford University Press papers
author:
  - name: Alice Anonymous
    email: alice@example.com
    corresponding_author: TRUE # Automatic footnote using corresponding_author
    affiliation:
      - ABC
      - DEF
    biography: |
      A promising young researcher working on her thesis project.
#    biography_pic: "/PATH/TO/image.png"
#    biography_pic_width: "1in"
  - name: Bob Security
    email: bob@example.com
    affiliation: DEF
    footnote: 1 # Footnote using reference
  - name: Cat Memes
    email: cat@anotheru.edu
    affiliation: Another University
    footnote:
        - 1
        - 2
address:
  - code: ABC
    department: Department of Biology
    organization: University of Idaho
    street: 123 Main Street
    #postcode: 
    city: Boise
    state: Idaho 07458 # Mix and match elements to get proper address
    country: United States
  - code: DEF
    address: Address in one element deprecated, but works
  - code: Another University
    address: Address for Another University
footnote:
  - code: 1
    text: "Equal contribution"
  - code: 2
    text: "Current email address: [cat@example.com](mailto:cat@example.com)"
abstract:
  - "This is the first paragraph of the abstract. It has several sentences making it go over several lines. For this it needs to have a lot of text." 
  - "This is the second paragraph."
  - "**Motivation:** You can also have some paragraphs start with bold face."
keywords:
  - keyword1
  - keyword2
  - keyword3
#boxedlist: # Not fully compatible with all document formats
#  - First point
#  - Second point
competing_interests: |
  There are no competing interest.
author_contributions: |
  AA did all the work. The others are just freeloaders.
acknowledgments: |
  This is an acknowledgement.

  It consists of two paragraphs.

bibliography: mybibfile.bib # Includes refs in OUP example template
## When using `citation_package = "default"`, a CSL file can be used
#csl: https://www.zotero.org/styles/oxford-university-press-note
output:
  rticles::oup_article:
    oup_version: 1 # 1 = 2020 CTAN OUP CLS package; 0 = 2009 OUP CLS package
    journal: "Journal Name"
    document_style: "contemporary" # Can be contemporary, modern, traditional
    papersize: "large" # Can be large, medium, small
    #citation_package: "default" # Uncomment when using a CSL; default "natbib"
    namedate: TRUE # Set FALSE to use numeric refs; Default FALSE
    #number_sections: FALSE # Uncomment to not number sections; default TRUE
    #number_lines: TRUE # Use `lineno` package - Default FALSE
    #number_lines_options: ["mathlines","switch"]  # Options for latex lineno package.
    #onecolumn: TRUE # Uncomment for one column format; default FALSE
    extra_dependencies:
      - booktabs # to use with knitr::kable() example below

## Example of pandoc's variable
#urlcolor: orange
#linkcolor: green
#citecolor: red
header-includes:
  #- \usepackage[nomarkers,tablesfirst]{endfloat} # For figures and tables at end
  - \theoremstyle{thmstyleone} # Theorem stuff from OUP template
  - \newtheorem{theorem}{Theorem} #  meant for continuous numbers. %%\newtheorem{theorem}{Theorem}[section] # meant for sectionwise numbers. optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
  - \newtheorem{proposition}[theorem]{Proposition} # %%\newtheorem{proposition}{Proposition}" # to get separate numbers for theorem and proposition etc.
  - \theoremstyle{thmstyletwo}
  - \newtheorem{example}{Example}
  - \newtheorem{remark}{Remark}
  - \theoremstyle{thmstylethree}
  - \newtheorem{definition}{Definition}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, # By default, hide code; set to TRUE to see code
  fig.pos = 'th', # Places figures at top or here
  out.width = '100%', dpi = 300, # Figure resolution and size
  fig.env="figure"
) # Latex figure environment

options(knitr.table.format = "latex") # For kable tables to write LaTeX table directly
```

# Introduction

This template is based on the generic OUP authoring template available on CTAN under [oup-authoring-template](https://www.ctan.org/pkg/oup-authoring-template). The CTAN template includes LaTeX documentation and a sample LaTeX document that provide far more details regarding the full functionality of the format. Here, only the basic functioning of the Rmarkdown adaptation of the format is demonstrated.

## A subsection

A numbered list:

1) First point
2) Second point
    - Subpoint

A bullet list:

* First point
* Second point

## Notes

* Extra white space in document will tend to disappear as text is filled in.
* Code blocks tend to generate lots of empty white space when `echo=TRUE` for some reason.

# Literature citations

By default, citations are handled by `natbib` using a numeric citation format. To use name-date citations, sets `namedate: TRUE` in the YAML header.

Here are two sample references:

* **author (year) example:** @horvath2018dna showed some really cool things. Only seems to work properly if `namedate: TRUE`. 
* **(author year) example:** This is a well known result [@ji20123d].

The bibliography will appear at the end of the document.

Though not normally available in the OUP LaTeX format, [CSL style files](https://www.zotero.org/styles) can also be used with the Rmarkdown adaptation by setting in the YAML header `citation_package: "default"` and defining the `csl` element to be the path towards the style file.

# Equations

An equation without a label for cross-referencing:

$$
E=mc^2
$$

An inline equation: $y=ax+b$

An equation with a label for cross-referencing:

\begin{equation}\label{eq:eq1}
\int^{r_2}_0 F(r,\varphi){\rm d}r\,{\rm d}\varphi = 1
\end{equation}

This equation can be referenced as follows: Eq. \ref{eq:eq1}

# Inserting R figures

The code below creates a figure. The code is included in the output because `echo=TRUE`.

```{r fig1, fig.cap="This is the first figure.",echo=TRUE}
plot(1:10,main="Some data",xlab="Distance (cm)",
     ylab="Time (hours)")
```

You can reference this figure as follows: Fig. \ref{fig:fig1}.

## Figures spanning two-columns

Figures can span two columns be setting `fig.env="figure*"`.

```{r fig2, fig.cap="This is a wide figure.",fig.env="figure*"}
plot(1:5,pch=19,main="Some wide data",
     xlab="Distance (cm)",ylab="Time (hours)")
```

Reference to second figure: Fig. \ref{fig:fig2}

# Tables

## Generate a table using `xtable`

```{r xtabletable, results="asis", echo=TRUE}
df = data.frame(ID=1:3,code=letters[1:3])

# Creates tables that follow OUP guidelines 
# using xtable
library(xtable) 
print(xtable(df,caption="This is a xtable table.",
             label="tab:tab1"),
      comment=FALSE,caption.placement="top")
```

You can reference this table as follows: Table \ref{tab:tab1}.

## Generate a table using `kable`

```{r kabletable, echo=TRUE}
df = data.frame(ID=1:3,code=letters[1:3])

# kable can alse be used for creating tables
knitr::kable(df,caption="This is a kable table.",
             booktabs=TRUE,label="tab2")
```

You can reference this table as follows: Table \ref{tab:tab2}.

## Table spanning two columns

Tables can span two columns be setting `table.envir = "table*"` in `knitr::kable`.

```{r widetable, echo=TRUE}
df = data.frame(ID=1:3,code1=letters[1:3],
                code2=letters[4:6],
                code3=letters[7:9],
                code4=letters[10:12],
                code5=letters[13:15])

# kable can alse be used for creating tables
knitr::kable(df,caption="This is a wide kable table.",
             #format="latex",
             table.envir="table*",
             booktabs=TRUE,label="tab3")
```

# Cross-referencing sections

You can cross-reference sections and subsections as follows: Section \ref{literature-citations} and Section \ref{a-subsection}.

***Note:*** the last section in the document will be used as the section title for the bibliography.

For more portable and flexible referencing of sections, equations, figures and tables, use [`bookdown::pdf_document2`](https://github.com/rstudio/bookdown) with YAML header option `base_format: rticles::oup_article`.

# Appendices {-}

::: {.appendices latex=true}

# Section title of first appendix

blabla

## Subsection title of first appendix

and so on....
:::

# References