### 2025-03-14 inst/misc/bmm_ppt.qmd
Add qmd ppt stuff

---
title: "hatchR"
format: pptx
editor: visual
execute:
  echo: TRUE
  warning: false
---

## Prep

```{r}
library(hatchR)
library(tidyverse)
library(ggridges) 
```

## Woody Island Dataset

```{r}
print(woody_island, n = 5)
```

## `plot_temp_check()`

```{r}
plot_check_temp(
  data = woody_island,
  dates = date, 
  temperature = temp_c
  ) +
  geom_rect(
    aes(
      xmin = ymd("1990-08-18"),
      xmax = ymd("1991-04-01"),
      ymin = -10,
      ymax = 25),
    fill = "grey",
    alpha = 0.01
    ) 
```

## Select built-in model parameterization

```{r}
model_table
```

------------------------------------------------------------------------

```{r}
sockeye_hatch_mod <- model_select(
  author = "Beacham and Murray 1990", 
  species = "sockeye", 
  model = 2, 
  development_type = "hatch"
  )
sockeye_hatch_mod$expression
```

## Predict phenology!

```{r}
WI_hatch <- predict_phenology(
  data = woody_island,
  dates = date,
  temperature = temp_c,
  spawn.date = "1990-08-18",
  model = sockeye_hatch_mod
  )
```

## Checking outputs

```{r}
WI_hatch$days_to_develop
WI_hatch$dev.period
```

## Plot phenology

```{r}
plot_phenology(WI_hatch)
```

------------------------------------------------------------------------

```{r echo=FALSE}
# Smallmouth Bass data (Webster 1948)
smallmouth <- matrix(NA, 10, 2) |> data.frame()
colnames(smallmouth) <- c("hours", "temp_F")
smallmouth$hours <- c(52, 54, 70, 78, 90, 98, 150, 167, 238, 234)
smallmouth$temp_F <- c(77, 75, 71, 70, 67, 65, 60, 59, 55, 55)

# change °F to °C and hours to days
smallmouth <- smallmouth |>
  mutate(
    days = ceiling(hours / 24),
    temp_C = (temp_F - 32) * (5 / 9)
  ) |> 
  select(-hours, -temp_F)
```

```{r}
smallmouth
```

------------------------------------------------------------------------

```{r}
# estimate parameters for Smallmouth Bass using fit_model()
smb_mod <- fit_model(
  temp = smallmouth$temp_C,
  days = smallmouth$days,
  species = "smb",
  development_type = "hatch"
)
```

------------------------------------------------------------------------

```{r}
smb_mod$pred_plot
```

------------------------------------------------------------------------

```{r}
smb_mod$r_squared
smb_mod$expression$expression
```

------------------------------------------------------------------------

```{r echo=FALSE, message=FALSE}
set.seed(123)

# create random temps and corresponding dates
temps_sim <- sort(rnorm(n = 30, mean = 16, sd = 1), decreasing = FALSE)
dates_sim <- seq(
  from = ymd("2000-07-01"),
  to = ymd("2000-07-31"), length.out = 30
)

data_sim <- matrix(NA, 30, 2) |> data.frame()
data_sim[, 1] <- temps_sim
data_sim[, 2] <- dates_sim

# change names so they aren't the same as the vector objects
colnames(data_sim) <- c("temp_sim", "date_sim")
```

```{r message=FALSE}
smb_hatch <- predict_phenology(
  data = data_sim,             # simulated data
  dates = date_sim,            # simulated dates
  temperature = temp_sim,      # simulated temperatures
  spawn.date = "2000-07-01",   # spawn date
  model = smb_mod$expression   # model expression from fit_model()
)
smb_hatch$days_to_develop
smb_hatch$dev.period
```

------------------------------------------------------------------------

```{r echo=FALSE}
# Summary for all species
all_res <- data.frame(matrix(NA, 1, 2))
colnames(all_res) <- c("start", "stop")

all_res$start <- c(smb_hatch$dev.period$start)

all_res$stop <- c(smb_hatch$dev.period$stop)

all_res <- all_res |>
  mutate(
    days = ceiling(stop - start),
    index = c(16.5)
  )

all_res$Species <- c("Smallmouth Bass")

ggplot() +
  geom_point(data = data_sim, aes(x = date_sim, y = temp_sim)) +
  geom_line(data = data_sim, aes(x = date_sim, y = temp_sim)) +
  geom_rect(data = all_res, aes(xmin = start, xmax = stop, ymax = index - .35, ymin = index - .5, fill = Species)) +
  geom_label(data = all_res, aes(x = start + (stop - start) / 1.25, y = (index - 0.425), label = days)) +
  labs(x = "Date", y = "Temperature (°C)") +
  scale_fill_manual(values = c("deepskyblue4", "grey23", "darkolivegreen4")) +
  theme_classic(base_size = 16) +
  theme(legend.position = c(0.75, 0.25))
```

------------------------------------------------------------------------

```{r}
plot_phenology(smb_hatch)
```

------------------------------------------------------------------------

```{r}
idaho
```

------------------------------------------------------------------------

```{r}
bull_trout_sites <- idaho |>
  mutate(month = month(date)) |>                # make a month column (numeric)
  filter(month == 8) |>                         # filter out Aug.
  group_by(site) |>                             # apply grouping by site
  summarise(mean_aug_temp = mean(temp_c)) |>
  filter(mean_aug_temp <= 13) |>                # keep only sites 13 C or cooler
  pull(site) |>
  unique()

# filter to bull trout sites
idaho_bt <- idaho |>
  filter(site %in% bull_trout_sites)
```

------------------------------------------------------------------------

```{r}
# nest data and summarize
isaak_summ_bt <- idaho_bt |>
  group_by(site) |>
  nest() |>
  mutate(
    summ_obj = map(
      data,
      summarize_temp,
      temperature = temp_c,
      dates = date
    )
  ) |>
  select(site, summ_obj)
```

------------------------------------------------------------------------

```{r}
isaak_summ_bt
```

------------------------------------------------------------------------

```{r}
# lets look at a couple individual sites
PIBO_1345 <- idaho_bt |> filter(site == "PIBO_1345")

# looks nice
plot_check_temp(PIBO_1345,
                dates = date,
                temperature = temp_c)
```

------------------------------------------------------------------------

```{r}
spawn_dates <- map(
  c(2011:2014),     # year vector to map for custom function
  function(year) {  # custom paste function
    c(
      paste0(year, "-09-01"),
      paste0(year, "-09-15"),
      paste0(year, "-09-30")
    )
  }
) |>
  unlist()
```

------------------------------------------------------------------------

```{r}
spawn_dates
```

------------------------------------------------------------------------

```{r}
bt_hatch <- model_select(
  development_type = "hatch",
  author = "Austin et al. 2019",
  species = "bull trout",
  model = "MM"
)
```

------------------------------------------------------------------------

```{r}
# apply predict_phenology() over multiple sites and spawn dates
hatch_res <- isaak_summ_bt |>
  mutate(
    dev_period = map2(
      summ_obj,                  # summarized temperature data object
      spawn_dates,               # list of spawn dates
      predict_phenology,         # function to apply
      temperature = daily_temp,  # column of temperature data in `summ_obj`
      dates = date,              # column of date in `summ_obj`
      model = bt_hatch           # model to use (from model_select())
    ) |>
      map_df("dev.period") |>    # pull out just dev.period results
      list()
  ) |>
  select(site, dev_period) |>           # just select the columns we want
  unnest(cols = c(dev_period)) |>       # un-nest everything
  mutate(days_to_hatch = stop - start)  # make a new column of days to hatch
```

------------------------------------------------------------------------

```{r}
cut_ints <- date(
  c(
    "2011-08-31", "2012-08-31",
    "2013-08-31", "2014-08-31",
    "2015-08-31"
  )
)

# add year and spawn time factors for plotting
hatch_factors <- hatch_res |>
  mutate(year = cut(start,
    breaks = cut_ints,
    labels = c(2011:2014)
  )) |>
  # cut coerces our labels to factors, this changes them back to numbers
  mutate(year = as.numeric(as.character(year))) |>
  mutate(day = day(start)) |>
  mutate(spawn_time = case_when(
    day == 1 ~ "Early",
    day == 15 ~ "Peak",
    day == 30 ~ "Late"
  )) |>
  mutate(spawn_time = factor(
    spawn_time,
    levels = c("Late", "Peak", "Early"),
    ordered = TRUE
  ))

hatch_factors_sum <- hatch_factors |>
  group_by(year, spawn_time, start) |>
  summarise(
    mean_hatch = ceiling(mean(days_to_hatch)),
    hatch_5 = ceiling(quantile(days_to_hatch, probs = 0.05)),
    hatch_95 = ceiling(quantile(days_to_hatch, probs = 0.95))
  )

ggplot(
  data = hatch_factors,
  aes(
    x = days_to_hatch,
    y = spawn_time,
    fill = spawn_time,
    color = spawn_time
  )
) +
  geom_density_ridges(alpha = 0.9) +
  facet_wrap(~year, ncol = 1) +
  scale_fill_brewer(palette = "Blues", direction = 1) +
  scale_color_brewer(palette = "Blues", direction = 1) +
  labs(x = "Days to hatch", y = "Spawn time") +
  theme_classic() +
  theme(legend.position = "none")
```